Teaching (Digital) Epigraphy

### 2016-01-20 source/48/EAGLE16-lame-et-al-teaching.tex
fixed few more papers
minor changes. THERE HAS BEEN CONFUSION WITH THE BIBLIO, MASTER AND
SOME OTHER FILES

A traditional print photograph, such as this one (No. 12582) drawn from the archives of the AMZ, limits the student’s
opportunity to learn to decipher these tags. Even if a professional photographer took this picture with a good quality
camera in a fully equipped photographic studio, the reading of the inscription stays riddled with uncertainties and a
researcher would have to check the tag de visu in order to offer a plausible transcription.

@@

From a digital epigraphy standpoint, this consists of an online DAP \ modeled, in Lamé 2015, on a dispositive analysis
and the three fundamental systems of an inscription: writing system (wSystem), textual system (tSystem) and contextual
system (cSystem). Three tools compose the DAP: RTI web viewer, TSS viewer and MarkOut tool. The last two were
specifically designed for the digital edition of the corpus TSS. The following paragraphs present those three DAP
tools. Due to the material properties of the lead tags, it is impossible to capture the appearance of the letters with
a single image. When holding a tag it is necessary to change the light angle to enhance the different marks on the
surface. RTI is a computational photography technique that enables the interactive re-lighting of the observed object
from any light direction. This is accomplished combining into a single compact data structure a set of fixed camera
photographs taken under many different illumination conditions. The acquisition process is inexpensive and does not
require labor intensive steps, unlike 3d model acquisition. RTI visualization allows the user to virtually ‘tilt’ the
object and recreate different lighting conditions, mimicking part of the autoptic process. Another advantage is the
magnification provided by the high resolution camera employed. The widespread adoption in browsers of WebGL (Khronos
group 2009) recently enable RTI visualization on the web. We used the RTI web viewer developed by Visual Computing Lab,
ISTI-CNR. High resolution re-illuminable images allow the selection of the best view for each mark, thus allowing the
user to obtain an adequate readability of the lead tags through an intuitive interface. Another important advantage of
having the tags available online is that the students can easily compare a tag with any other tag already in the
database and compare occurrences of a letter in other already transcribed tags. Finally, the Web access facilitates the
discussion amongst students and experts worldwide.

@@

Fig. \ref{fig:markout}

>>>>

A traditional print photograph, such as this one (No. 12582, Fig.~\ref{fig:12582}) drawn from the archives of the AMZ, limits the student’s
opportunity to learn to decipher these tags. Even if a professional photographer took this picture with a good quality
camera in a fully equipped photographic studio, the reading of the inscription stays riddled with uncertainties and a
researcher would have to check the tag \emph{de visu} in order to offer a plausible transcription.

@@

From a digital epigraphy standpoint, this consists of an online DAP \ modeled, in \citet{Lame2015}, on a dispositive analysis
and the three fundamental systems of an inscription: writing system (wSystem), textual system (tSystem) and contextual
system (cSystem). Three tools compose the DAP: RTI web viewer, TSS viewer and MarkOut tool. The last two were
specifically designed for the digital edition of the corpus TSS. The following paragraphs present those three DAP
tools. Due to the material properties of the lead tags, it is impossible to capture the appearance of the letters with
a single image. When holding a tag it is necessary to change the light angle to enhance the different marks on the
surface. RTI is a computational photography technique that enables the interactive re-lighting of the observed object
from any light direction. This is accomplished combining into a single compact data structure a set of fixed camera
photographs taken under many different illumination conditions. The acquisition process is inexpensive and does not
require labor intensive steps, unlike 3d model acquisition. RTI visualization allows the user to virtually ‘tilt’ the
object and recreate different lighting conditions, mimicking part of the autoptic process. Another advantage is the
magnification provided by the high resolution camera employed. The widespread adoption in browsers of WebGL (Khronos
group 2009) recently enable RTI visualization on the web. We used the RTI web viewer developed by Visual Computing Lab,
ISTI-CNR. High resolution re-illuminable images allow the selection of the best view for each mark, thus allowing the
user to obtain an adequate readability of the lead tags through an intuitive interface. Another important advantage of
having the tags available online is that the students can easily compare a tag with any other tag already in the
database and compare occurrences of a letter in other already transcribed tags. Finally, the Web access facilitates the
discussion amongst students and experts worldwide.

@@
